### 2020-05-06 thesis_template/2_RelatedWork/model/gctuning.tex
need object size and conclusion documentation

\subsection{Resilient Distributed Datasets}
Major methods in Spark are Resilient Distributed Datasets (RDDs), a data structure that abstracts distributed memory across different clusters. 
The immutable coarse-grained transformation, spark-scheduler with lazy-evaluation, and memory management with cacheing achieve computation with fault-tolerance, 
fast execution, and moderate control on memory efficiency \cite{DBLP:conf/nsdi/ZahariaCDDMMFSS12}.

A RDD is essentially a multi-layer Java data structure. A top RDD object references Java array, which intern, references a set of tuple objects. 
The coarse-grained transformations and immutability requires a RDD to be deep-copied to produce a new RDD, but efficiently offers fault tolerance. 
The lost partitions of a RDD can be recomputed in parallel on different nodes rather than rolling back the whole program.

Spark-pipeline consists of sequence of transformations and actions over RDDs. A transformation produces a new RDD from a set of existing RDD. 
An action is method that computes statistics from an RDD. Due to lazy-evaluation nature, transformations do not materialize the newly created RDD. 
Instead, RDD Lineages are created. Lineage is a graph among parent and child RDDs which represents logical execution plan. 
This enhance fault-tolerance and improve ability to optimize execution plan.

RDDs can be cached in memory for faster access by persist method. Developers can specify a storage level for a persisted RDD, in memory with serialized or deserialized, or on disk. 
Other than persisted RDD, Spark generates a lot of intermediate RDDs during execution. Since RDD is a Java object, they are managed by Garbage Collection (GC) in the JVM. 
However, persisted RDDs are never collected by GC. This GC might cause significant deterioration of performance of Spark, because GC shows heavy overhead when there are a number of objects.

\subsection{Memory Management in Spark}
Spark framework allocates multiple executors, JVMs, that run sequence of transformations and actions. As we describe the previous section, data in Spark is mainly stored as Java objects in memory, 
so that they are allocated on JVM heap and managed by JVM Garbage Collection. The data may form three types \cite{DBLP:journals/pvldb/XuGDWW19} \cite{DBLP:conf/pldi/WangCCZVML0X19}: Cached data, Shuffled data, and Operator-generated data.

Spark can cache data in memory to reduce disk I/O. This Cached data usually long-lived Java objects and span multiple stages in Spark-pipeline. 
Spark allocates a logical storage space to store the cached data as shown in Figure. After aggregation, Spark generates Shuffled data. Shuffled data is usually long-lived, 
because it need to be kept in memory until the task ends. Spark allocates execution space to store Shuffled data.
The storage space and execution space spans 60 \% of JVM heap space in default. Operator-generated data is data generated by user-defined operations. Since Operator-generated data may or may not be used, 
after the operation finish, the data object can be both short-lived or long-lived objects. These are stored in user space allocated on default 40 \% of JVM heap.

All data of these types on JVM heap is managed by JVM GC. GC check references graph of objects, mark whether the objects are used and deallocate memory space occupied by unused objects. 
There are three popular GCs: Parallel, CMS and G1. All of these method track generation of object based on the region of memory. 
The Java logical heap structure is shown in Figure~\ref{fig:javaheap}. As the figure shows, Java logical heap can be separated into three main parts where store objects for each corresponding generation: 
permanent generation, young generation, and old generation. The region for permanent generation stores metadata required by JVM to describe 
class and method used in application which will be permanently lived on the region of memory. The overview of Java GC is shown in Figure~\ref{fig:javagc}. 
The outer box represents region for particular generation. The inner box and the number represents object and its age in GC.

The region for young generation mainly consists of two parts: Eden and Survivor space. 
First, Java objects are created in Eden space and promoted to Survivor space when survive from GC. 
After objects survive several GCs in Survivor space, they are finally promoted to old generation.

In region of old generation, JVM lunches multi-thread to perform GC. GC with multi-threading suffers from Stop-The-World (STW) pauses; 
GC may suspend application threads while performing object marking and deallocation. 
Different GC algorithms try to solve this problem with trade-off between GC frequency and memory utilization.

Because of the problem of STW and copying objects to different physical memory pages , JVM GC cause huge overhead when number of objects is large. 
Therefore, GC become severe issue in Big Data processing where might produce significant number of object.

\begin{figure}[htb]
    \includegraphics[width=15cm]{java_heap.eps}
    \caption{Java Heap Structure}
    \label{fig:javaheap}
\end{figure}

\begin{figure}[htb]
    \includegraphics[width=15cm]{java_gc.eps}
    \caption{Java Garbage Collection}
    \label{fig:javagc}
\end{figure}

\subsection{Garbage Collection Tuning}
There are many different ways that one can improve the performance of GC in Java. One of these is for example avoiding pointer-based data structures, such as HashMap and LinkedList. 
These objects have a "wrapper" object for each entry so that number of object tends to be larger than when when an array is used.

>>>>

There are many different ways that one can improve the performance of GC in Java. One of these is for example avoiding pointer-based data structures, such as HashMap and LinkedList. 
These objects have a "wrapper" object for each entry so that number of object tends to be larger than when when an array is used.

### 2020-05-06 thesis_template/2_RelatedWork/model/linalg.tex
need object size and conclusion documentation

Basic Linear Algebra Subprograms (BLAS) are standard building blocks for basic vector and matrix operations. There are 3 levels of operation. The level 1 BLAS performs scalar, 
vector and vector-vector operations, the level 2 BLAS performs matrix-vector operation, and the level 3 BLAS performs matrix-matrix operation.

LAPACK is developed on BLAS and has advanced functionalities such as LU decomposition and Singular Value Decomposition (SVD). Dense and banded matrices are handled, but not general 
sparse matrices. The initial motivation of development of LAPACK was to make the widely used EISPACK and LINPACK libraries run efficiently on shared-memory vector and parallel processors. 
LINPACK and EISPACK are inefficient because their memory access patterns disregard the multi-layered memory hierarchies of the machines, thereby spending too much time moving data. 
LAPACK addresses this problem by recognizing the algorithms to use block matrix operations, such as matrix multiplication, in the innermost loops. These block operations can be optimized for each architecture to account for the memory hierarchy, 
and so provide a portable way to achieve high efficiency on diverse modern machines. However, LAPACK requires that highly optimized block matrix operations be already implemented on each machine.

ARPACK is also a collection of linear algebra subroutines which is designed to compute a few eigenvalues and corresponding eigenvectors from large scale matrix. 
ARPACK is based upon an algorithmic variant of the Arnoldi process called the Implicitly Restarted Arnoldi Method (IRAM). 
When the matrix is symmetric it reduces to a variant of the Lanczos process called the Implicitly Restarted Lanczos Method(IRLM). The Arnoldi process only interacts with the matrix via matrix-vector multiplies. 
Therefore, this method can be applied to distributed matrix operations required in big data analysis.

\subsection{Netlib-Java}
\label{sec:linalg_netlib}

Netlib-java is a Java wrapper of BLAS, LAPACK, and ARPACK. Netlib-java choose  implementation of linear algebra depending on installation of the libraries. 
First, if we have installed machine optimized system libraries, such as Intel MKL and OpenBLAS, netlib-java will use these as the implementation to use. 
Next, it try to load netlib references which netlib-java use CBLAS and LAPCKE interface to perform BLAS and LAPACK native call. 
The last option is to use f2j which is intended to translate the BLAS and LAPACK libraries from their Fortran77 reference source code to Java class files, 
instead of calling native libraries by using Java Native Interface (JNI).

>>>>

Basic Linear Algebra Subprograms (BLAS) \cite{DBLP:journals/toms/LawsonHKK79} are standard building blocks for basic vector and matrix operations. There are 3 levels of operation. The level 1 BLAS performs scalar, 
vector and vector-vector operations, the level 2 BLAS performs matrix-vector operation, and the level 3 BLAS performs matrix-matrix operation.

LAPACK \cite{laug} is developed on BLAS and has advanced functionalities such as LU decomposition and Singular Value Decomposition (SVD). Dense and banded matrices are handled, but not general 
sparse matrices. The initial motivation of development of LAPACK was to make the widely used EISPACK \cite{DBLP:books/sp/SmithB76} and LINPACK \cite{DBLP:journals/sigarch/Dongarra92} libraries run efficiently on shared-memory vector and parallel processors. 
LINPACK and EISPACK are inefficient because their memory access patterns disregard the multi-layered memory hierarchies of the machines, thereby spending too much time moving data. 
LAPACK addresses this problem by recognizing the algorithms to use block matrix operations, such as matrix multiplication, in the innermost loops. These block operations can be optimized for each architecture to account for the memory hierarchy, 
and so provide a portable way to achieve high efficiency on diverse modern machines. However, LAPACK requires that highly optimized block matrix operations be already implemented on each machine.

ARPACK \cite{DBLP:books/daglib/0000896} is also a collection of linear algebra subroutines which is designed to compute a few eigenvalues and corresponding eigenvectors from large scale matrix. 
ARPACK is based upon an algorithmic variant of the Arnoldi process called the Implicitly Restarted Arnoldi Method (IRAM). 
When the matrix is symmetric it reduces to a variant of the Lanczos process called the Implicitly Restarted Lanczos Method(IRLM). The Arnoldi process only interacts with the matrix via matrix-vector multiplies. 
Therefore, this method can be applied to distributed matrix operations required in big data analysis.

\subsection{Netlib-Java}
\label{sec:linalg_netlib}
Netlib-java \cite{netlib} is a Java wrapper of BLAS, LAPACK, and ARPACK. Netlib-java choose  implementation of linear algebra depending on installation of the libraries. 
First, if we have installed machine optimized system libraries, such as Intel MKL and OpenBLAS, netlib-java will use these as the implementation to use. 
Next, it try to load netlib references which netlib-java use CBLAS and LAPCKE interface to perform BLAS and LAPACK native call. 
The last option is to use f2j which is intended to translate the BLAS and LAPACK libraries from their Fortran77 reference source code to Java class files, 
instead of calling native libraries by using Java Native Interface (JNI).

### 2020-05-06 thesis_template/2_RelatedWork/model/spark.tex
ready to review
need object size and conclusion documentation

\subsection{Memory Management in Spark}
Spark framework allocates multiple executors, JVMs, that run sequence of transformations and actions. As we describe the previous section, data in Spark is mainly stored as Java objects in memory, 
so that they are allocated on JVM heap and managed by JVM Garbage Collection. The data may form three types \cite{DBLP:journals/pvldb/XuGDWW19}: Cached data, Shuffled data, and Operator-generated data.

@@

\subsection{Garbage Collection Tuning}
There are many different ways that one can improve the performance of GC in Java. One of these is for example avoiding pointer-based data structures, such as HashMap and LinkedList. 
These objects have a "wrapper" object for each entry so that number of object tends to be larger than when when an array is used.

Caching serialized object in memory also reduce number of object and memory usage since the set of objects become a byte or binary array. 
Spark SQL applications use DataFrames\cite{DBLP:conf/sigmod/ArmbrustXLHLBMK15}, 
whose intermediate data are managed by an optimized memory manager named Tungsten.
Tungsten stores the intermediate data in a serialized binary form and performs aggregation functions directly on he serialized objects.
Therefore, the number of Java objects in memory is reduced and it reduces the DC frequency and object marking/sweeping.

In addition, developers can allocate data off-heap of JVM to avoid tracking by GC. 
Facade\cite{DBLP:conf/asplos/NguyenWBFHX15} proposed a compiler and runtime system to bound the number of in-memory data objects, through storing data in an off-heap region and 
manipulating the data with control interfaces.

Although these memory management solution for GC help developer improve performance of Spark applications, 
the effort to discover the best GC tuning afflicts developers.

>>>>

\subsection{Memory Management in Spark}
Spark framework allocates multiple executors, JVMs, that run sequence of transformations and actions. As we describe the previous section, data in Spark is mainly stored as Java objects in memory, 
so that they are allocated on JVM heap and managed by JVM Garbage Collection. The data may form three types \cite{DBLP:journals/pvldb/XuGDWW19} \cite{DBLP:conf/pldi/WangCCZVML0X19}: Cached data, Shuffled data, and Operator-generated data.

@@

### 2020-05-06 thesis_template/3_Body/chapter/evaluation.tex
done
revise untill intro
need object size and conclusion documentation

\section{Introduction}
\label{sec:eval_intro}
In this chapter, detail setting, result, and analysis of our experiments are discussed.
Section~\ref{sec:eval_setdetail} details hardware configurations and data set we use. 
There are 5 experiments conducted. In Section~\ref{sec:eval_diffval}, we discuss the experiment 
where we assess runtime to access object's field with different variable types. 
Section~\ref{sec:eval_refcount} shows details to examine how behaviors of normal reference and Reference Count are different. 
Section~\ref{sec:eval_sort} describes our Merge-sort experiment to examine behavior of Atomic Reference Counting.
Two algorithms are discussed used in Big Data processing and examined in our experiments: 
Tree-aggregate in Section~\ref{sec:concept_treeagg} and K-Nearest-Neighbors in Section~\ref{sec:concept_knn}.

@@

\section{Experiment 2: Merge-sort}
\label{sec:eval_sort}
\input{3_Body/model/merge_sort_experiment}

@@

\section{Summary}
\label{sec:eval_summary}
In this chapter, we examines various memory management strategies for different algorithms in Rust programming.
There are many factors that might differ performance of algorithms: different variable type, Reference Counting, Atomic Reference Counting, and frequency of memory de/allocation.
We should select the best memory management strategies when developing Big Data processing tools. 
Some of tips for the development are proved in our experiments. We conclude this dissertation in Chapter~\ref{chapter:Conclusions}.

>>>>

\section{Introduction}
\label{sec:eval_intro}
In this chapter, detail experiment settings, executions and results are discussed.
Section~\ref{sec:eval_setdetail} details hardware configurations and data set we use. 
There are 5 experiments conducted. In Section~\ref{sec:eval_diffval}, we discuss the experiment 
where we assess runtime to access object's field with different variable types. 
Section~\ref{sec:eval_refcount} shows details to examine how behaviors of normal reference and Reference Count are different. 
Section~\ref{sec:eval_sort} describes our Merge-sort experiment to examine behavior of Atomic Reference Counting.
Two algorithms are discussed used in Big Data processing and examined in our experiments: 
Tree-aggregate in Section~\ref{sec:concept_treeagg} and K-Nearest-Neighbors in Section~\ref{sec:concept_knn}.

@@

\section{Experiment 3: Merge-sort}
\label{sec:eval_sort}
\input{3_Body/model/merge_sort_experiment}

@@

\section{Summary}
\label{sec:eval_summary}
In this chapter, we examines various memory management strategies for different algorithms in Rust programming.
There are many factors that might differ performance of algorithms: different variable type, Reference Counting, Atomic Reference Counting, and frequency of memory de/allocation.
We should select the best memory management strategies when developing Big Data processing tools. 
Some of tips for the development are proved in our experiments. We conclude this thesis in Chapter~\ref{chapter:Conclusions}.

### 2020-05-06 thesis_template/3_Body/model/experimental_set_detail.tex
need object size and conclusion documentation

\subsection{Wikipedia Data Sets}
 \label{sec:concept_dataset}
 Wikipedia page data sets are used to perform document classification with KNN. 
 \(10^5\) pages are used for training data set, and 18724 pages are used for target.

 \subsection{Experimental Details}
 \label{sec:concept_expdetail}
 Our experiments are run on VM instances on Google Cloud Platform, 
 n1-standard-8 which has 8 vCPU, 30 GB RAM, and 10 GB Standard persistent disk.
 In this thesis, we present the average of 5 separate runs for each experiment.

>>>>

We have used two different data sets. One is a real-world data set, Wikipedia page data set,
and another is a syntactic randomized generated data set based on Complex object described in Section~\ref{sec:concept_compobj}. 
We use one machine with particular specification.
\subsection{Wikipedia Data Sets}
 \label{sec:concept_dataset}
 Wikipedia page data sets are used to perform document classification with KNN. 
 We have separated the data set into training and test data set, 
 \(10^5\) pages are used for training data set, and 18724 pages are used for target.

 \subsection{Experimental Details}
 \label{sec:concept_expdetail}
 All experiments are run on VM instances on Google Cloud Platform, 
 n1-standard-8 which has 8 vCPU, 30 GB RAM, and 10 GB Standard persistent disk.
 In this thesis, we present the result of runtime as the average of 5 separate runs for each experiment.

### 2020-05-06 thesis_template/3_Body/model/knn.tex
need object size and conclusion documentation

The algorithms have 4 phase; load, preprocess, query, and combine phase. 
The algorithms spawn threads at the beginning ,and in the load phase partition of files are loaded in each threads. 
In preprocessing phase, the algorithm process document strings to generate Term-frequencies (Tfs) matrices and other data structures. 
In query phase, it calculates cosine similarities between all combination of train and test observations and select to top \(K\) nearest neighbors. 
In combine phase, the results from each batch and from each are combined. 
Based on our experiments result, the runtime in preprocess and query phase are significantly larger than other two phases.
Therefore, we focus discussion in these two phases considering them bottlenecks in our KNN algorithms.

@@

\begin{table}
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{|l|l|l|l|l|}
    \hline
    Parameter Name  & \multicolumn{4}{l|}{Values and Description}                                                                                                                                                                                \\ \hline
    Method          & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}deep-copy:  use deep-copy to generate intermediate objects\\ arc: use atomic reference count to generate intermediate objects\end{tabular}}                                 \\ \hline
    Strategy        & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}1: keep intermediate objects in memory until owner is changed\\ 2: remove intermediate objects as soon as it is not needed\end{tabular}}                                    \\ \hline
    Number of batch & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}2: generate 2 batch from each partition\\ 3: generate 3 batch from each partition\end{tabular}}                                                                             \\ \hline
    k               & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}15K: dimension of feature matrices is 15 thousands \\ 20K: dimension of feature matrices is 20 thousands\\ 25K: dimension of feature matrices is 25 thousands\end{tabular}} \\ \hline
    \end{tabular}
    \caption{Parameter of KNN algorithms}
    \label{tab:parameter}
 \end{table}

@@

Number of batch controls number and size of batch for each partition. This parameter varies size of intermediate data structures and numeric matrices, and 
frequency of memory deallocations. 
The parameter k specifies number of dimension of feature matrices created out of preprocess phase. 
By controlling this, it determines size of intermediate data structures and numeric matrices.

@@

The impact of number of batch can explain how frequency of memory de/allocation affects runtime performance. 
Dealing with more number of batch has negative impacts to runtime performance, because it triggers more frequent memory de/allocation.
In addition, as we increase the number of dimension, the ration of runtime difference among different numbers of batch for the algorithms with deep-copy, but not for ones with Arc.
This is because algorithm with Arc does not de/allocate memory as much frequent as ones with deep-copy.
The intermediate data structures mealy get reference to String values without actually allocating memory for copied Strings.

>>>>

The algorithms have 4 phase; load, preprocess, query, and combine phase. 
Before these phase, we separate data sets into 8 partitions to run these in different threads. 
The algorithms spawn threads at the beginning ,and in the load phase partition of files are loaded in each threads. 
In preprocessing phase, the algorithm process document strings to generate Term-frequencies (Tfs) matrices and other data structures. 
In query phase, it calculates cosine similarities between all combination of train and test observations and select to top \(K\) nearest neighbors. 
In combine phase, the results from each batch and from each are combined. 
Based on our experiments result, the runtime in preprocess and query phase are significantly larger than other two phases.
Therefore, we focus discussion in these two phases considering them bottlenecks in our KNN algorithms.

@@

\begin{table}
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{|l|l|l|l|l|}
    \hline
    Parameter Name  & \multicolumn{4}{l|}{Values and Description}                                                                                                                                                                                \\ \hline
    Method          & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}deep-copy:  use deep-copy to generate intermediate objects\\ arc: use atomic reference count to generate intermediate objects\end{tabular}}                                 \\ \hline
    Strategy        & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}1: keep intermediate objects in memory until owner is changed\\ 2: remove intermediate objects as soon as it is not needed\end{tabular}}                                    \\ \hline
    Number of batch & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}2: generate 2 batches from each partition\\ 3: generate 3 batches from each partition\end{tabular}}                                                                             \\ \hline
    k               & \multicolumn{4}{l|}{\begin{tabular}[c]{@{}l@{}}15K: dimension of feature matrices is 15 thousands \\ 20K: dimension of feature matrices is 20 thousands\\ 25K: dimension of feature matrices is 25 thousands\end{tabular}} \\ \hline
    \end{tabular}
    \caption{Parameter of KNN algorithms}
    \label{tab:parameter}
 \end{table}

@@

Number of batch controls number and size of batch for each partition. This parameter varies size of intermediate data structures and numeric matrices, and 
frequency of memory deallocations. 
The size of each batch is defined in 6250 pages when we have 2 batches, and 4166 or 4167 when we have 3 batches. 
The parameter k specifies number of dimension of feature matrices created out of preprocess phase. 
By controlling this, it determines size of intermediate data structures and numeric matrices.

@@

The impact of batch number can explain how frequency of memory de/allocation affects runtime performance. 
Dealing with more number of batch has negative impacts on runtime performance because it triggers more frequent memory de/allocation.
In addition, as we increase the number of dimension, the ration of runtime difference among different numbers of batch for the algorithms with deep-copy, but not for ones with Arc.
This is because algorithm with Arc does not de/allocate memory as much frequent as ones with deep-copy.
The intermediate data structures mealy get reference to String values without actually allocating memory for copied Strings.

### 2020-05-06 thesis_template/3_Body/model/merge_sort_experiment.tex
need object size and conclusion documentation

As explained in last experiment, deletion of Rc has significant overhead than normal reference. 
By learning this result, our assumption here is that deletion of Arc has also overhead when we compare to normal reference. 
In many situation where developer writes a multithreading code, Arc is used and the deletion happens multiple times. 
To assess runtime overhead of algorithm with Arc, We implement merge-sort algorithm in two different ways. 
One is sharing source vector with Arc. The other is passing reference of source vector to child thread.

>>>>

Sorting algorithm like merge-sort is a very ubiquitous algorithm in any computation. 
Especially in Big Data processing, merge-sort algorithm de/allocates contiguous memory for partitions of sorted vector multiple times.
We are interested in memory de/allocation pattern in merge-sort algorithm, because it generates many intermediate data structures and delete them many times.

As explained in Experiment 2 in Section~\ref{sec:eval_refcount}, deletion of Rc has significant overhead than normal reference. 
By learning this result, our assumption here is that deletion of Arc has also overhead when we compare to normal reference. 
In many situation where developer writes a multithreading code, Arc is used and the deletion happens multiple times. 
To assess runtime overhead of algorithm with Arc, We implement merge-sort algorithm in two different ways. 
One is sharing source vector with Arc. The other is passing reference of source vector to child thread.
